### 2022-10-20 index.Rmd
Up

---
title: "Gov 50 Final Project"
author: "Mashal Farrukh"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Project thoughts

I am interested in exploring data related to the Covid 19 Pandemic and the efficacy of Pharmaceutical and Non-Pharmaceutical interventions (lock downs, contact tracing etc.) in controlling the spread of the virus globally.