### 2022-10-20 index.Rmd
updated

---
title: "Gov 50 Final Project"
author: "Mashal Farrukh"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Project thoughts

I am interested in exploring data related to the Covid 19 Pandemic and the efficacy of Pharmaceutical and Non-Pharmaceutical interventions (lock downs, contact tracing etc.) in controlling the spread of the virus globally. I want to compare the effectiveness of these across different social groups by using categories like gender, country etc.

I am also interested in exploring data related to the Covid 19 Pandemic and the how that has impacted Gender equity, especially domestic violence against women in different countries.