### 2022-10-20 article.Rmd
hh

---
title: "Gov 50 Final Project"
description: |

I am interested in exploring data related to the Covid 19 Pandemic and the efficacy of Pharmaceutical and Non-Pharmaceutical interventions (lock downs, contact tracing etc.) in controlling the spread of the virus globally. I want to compare the effectiveness of these across different social groups by using categories like gender, country etc.

I am also interested in exploring data related to the Covid 19 Pandemic and the how that has impacted Gender equity, especially domestic violence against women in different countries. 
author:
  - name: Mashal Farrukh 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.