### 2022-12-04 article.Rmd
j

---
title: "Gov 50 Final Project"
description: |

author:
  - name: Mashal Farrukh 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Project thoughts

I am interested in exploring the relationship between religious practice, party affliations, on people's attitude towards COVID restrictions.

```{r}
library(tidyverse)
library(readr)
```

```{r}
ATP_W70 <- read.csv("ATP_W70FromR.csv")
library(tidyverse)
```

```{r}
ATP_W70 <- ATP_W70 |> mutate (Pray = case_when(SPIRIT2_a_W70 >=1 & SPIRIT2_a_W70 <= 4 ~ '1')) |> mutate(Scripture = case_when(SPIRIT2_b_W70 >=1 & SPIRIT2_b_W70 <= 4 ~ '1')) |> mutate (Meditate = case_when(SPIRIT1_a_W70 >=1 & SPIRIT1_a_W70 <= 4 ~ '1')) |> mutate (spirtuality = ifelse (Pray == 1| Scripture == 1| Meditate == 1, "1", "0"))

```

```{r}
ATP_W70 <- ATP_W70 |> mutate(Opinions_on_SD = case_when(COVIDCONG4_W70 == 1 ~ "Open", COVIDCONG4_W70 == 2 ~ "Open with restictions", COVIDCONG4_W70 == 3 ~ "Closed"))|> mutate (college_edu = ifelse(F_EDUCCAT == 1, "College", "Not College")) |> mutate (PARTY_affiliation = case_when(PARTY_W70 == 1 ~ "Republican", PARTY_W70 == 2 ~ "Democrat", PARTY_W70 == 3 ~ "Independent", PARTY_W70 == 4 ~ "Something else")) |> mutate (Black = if_else(F_RACECMB == 2, "1", "0")) |> mutate(Ideology = case_when(F_IDEO > 0 & F_IDEO < 3 ~ "Conservative", F_IDEO == 3 ~ "Moderate", F_IDEO > 3 & F_IDEO < 5 ~ "Liberal"))

Opinions_final <- ATP_W70 |> drop_na(Opinions_on_SD) |> select(Opinions_on_SD, COVIDCONG4_W70, college_edu, PARTY_affiliation, Black, F_IDEO, Ideology, spirtuality) |> filter (PARTY_affiliation %in% c("Republican", "Democrat"))

Opinions_final  
```

##visualization 1
```{r}
Opinions <- Opinions_final |> group_by( PARTY_affiliation, Opinions_on_SD) |> summarise(n= n()) |> mutate (prop = n/ sum(n)) |> drop_na( PARTY_affiliation, Opinions_on_SD)

Opinions_plot <- Opinions |> ggplot(aes(x = Opinions_on_SD, y = prop, fill = PARTY_affiliation)) + geom_col(position = "dodge") + labs( Title =  "opinions on what the operatings status of places of worship should be according to Party affiliation")

Opinions_plot
```
Visualisation # 2
```{r}

Opinions_diff_by_edu <- Opinions_final |> filter (PARTY_affiliation %in% c("Democrat", "Republican")) |> group_by(PARTY_affiliation, college_edu, Opinions_on_SD) |> summarise(n= n()) |> mutate (prop = n/ sum(n)) |> pivot_wider(names_from = PARTY_affiliation, values_from = prop)

#|> drop_na(Opinions_on_SD, PARTY_affiliation) 
knitr:: kable (Opinions_diff_by_edu)

Opinions_by_educ_plot <- Opinions_diff_by_edu|> ggplot(aes(x = Opinions_on_SD, y = Democrat, fill =  college_edu)) + geom_col(position = "dodge") + labs( Title = "opinions on what the operatings status of places of worship should be according to Party affiliation and college education")

Opinions_by_educ_plot
```
```{r}

Opinions_diff_by_edu_rep <- Opinions_final |> filter (PARTY_affiliation %in% c("Democrat", "Republican")) |> group_by(PARTY_affiliation, college_edu, Opinions_on_SD) |> summarise(n= n()) |> mutate (prop = n/ sum(n)) |> select (-n)|> pivot_wider(names_from = PARTY_affiliation, values_from = prop)

#|> drop_na(Opinions_on_SD, PARTY_affiliation) 
knitr:: kable (Opinions_diff_by_edu_rep)

Opinions_by_educ_plot_rep <- Opinions_diff_by_edu_rep|> ggplot(aes(x = Opinions_on_SD, y = Republican, fill =  college_edu)) + geom_col(position = "dodge") + labs( Title = "opinions on what the operatings status of places of worship should be according to Party affiliation and college education")

Opinions_by_educ_plot_rep
```
```{r}
ATP_W70_TEST |> mutate(Opinions_on_SD_numeric = case_when(COVIDCONG4_W70 == 1 ~ "1", COVIDCONG4_W70 == 2 ~ "0", COVIDCONG4_W70 == 3 ~ "-1")) |> mutate (Party_binary = case_when(PARTY_W70 == 1 ~ "1", PARTY_W70 == 2 ~ "0")) |> select (Party_binary, PARTY_W70) 
```

```{r}
library(tidyverse)
ATP_W70_TEST <- ATP_W70 |> drop_na(COVIDCONG4_W70)

sing_Regression <- lm(Opinions_on_SD_numeric ~ party_binary, data = ATP_W70_TEST)
summary(sing_Regression)

```

```{r}
Mult_regression <- lm(Opinions_on_SD ~ party-binary + Black + college_edu + spirtuality, data = ATP_W70)
```

My research question aims to explore whether Party affiliations have any impact on people's opinions on the operating status of their houses of worship. For this, I am looking at the responses of those people who have a tendency to attend religious services. My main explanatory variable is Party affiliations and my outcome variable is opinions on social distancing (Opinions_on_SD). Since this is a cross-sectional research design and I cannot randomize my experiment fully, I want to control for the factors that could possibly confound my analysis. Some of these include education levels of the respondents, the race, as well as inherent spirituality.

For this milestone I added the education variable to my analysis to see if having completed college makes an impact on the difference in proportions on the opinions on social distancing between republicans and democrats. The visualization and the table show that having attained college education does not drastically change the trend that we observed in the first visualization in the sense that on average, the democrats only exceed the republicans as far as opinions on places of worship being closed is concerned. For both the other options, the proportions of republicans is greater than than of democrats. Adding the education variables enables us to see how the differences in proportion varies across college graduates and non-graduates and adds more precision to our analysis.